Thesis: draft

### 2021-04-21 draft.tex
continued writing results section

As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions. On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network. Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\ 
\newline
In one of their previous papers (\textit{"Modelling opinion dynamics in the age of algorithmic personalization"}, see reference \cite{Perra2019}) prof. Nicola Perra and prof. Luis E.C. Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions. They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).\\ Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}. On the other hand, a more heterogeneous contact pattern could hamper this formation of echo chambers \cite{Perra2019}. If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}. In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.\\
They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure. This thesis will build on their model and incorporate these new features. We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}). We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can, for example, observe convergence to one opinion inside the communities. We will also pay some attention to networks that have both a high clustering coefficient and a community structure. Can we observe echo chambers in those kinds of networks? If yes, can we determine which property might be the main driver for the formation of echo chambers: high clustering or community structure? \\
% ALSO mention real-world network??
\newline
On the other hand, we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network. Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness? What about the possible interplay between the filtering algorithms and the introduction of stubborn people? Does the distribution of stubborn people play a role (eg. randomly distributed versus clustered together inside communities)? \\

@@

Lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$. In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted. The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$). This is a measure for the formation of echo chambers (opinion clusters) in the system. The data points represent the value of the echo chamber size in networks with a different modularity. The initial opinion distribution was always $50/50$. \\

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_PR_20-80_8x7.png}
    \caption{PR}
    \label{PR_all_frac_stubb_20-80}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_REC_20-80_8x7.png}
    \caption{REC}
    \label{REC_all_frac_stubb_20-80}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution. \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_all_frac_stubb_PR-REC_20-80}
\end{figure}
\noindent
We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn. Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter r ($r=0$, $r=0.4$ and $r=0.8$). Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
\newline
Figure \ref{ev_op_20_80_all_frac_stubb_REC} gives the results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm. For none of the network models, nor the values of the stubbornness parameter $r$, the initial prevalence of the opinions can be broken. This result is analogue to the result for an initial $20/80$ opinion distribution, combined with the REC filtering algorithm, in the case of introducing a fraction of completely stubborn people (see Figure \ref{ev_op_20_80_stubb_REC}).

@@

In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion 0. The remaining communities are given a $50/50$ opinion distribution. This situation may, in an idealized way, represent the situation in which you have groups of people following the same idea together with groups where the two ideas coexist. The total opinion distribution over the entire network is no longer $50/50$, but instead opinion 0 becomes the majority opinion (hence, opinion 1 becomes the minority opinion). The higher the fraction of single opinion 0 communities, the bigger the difference between this majority and minority opinion becomes. The results will always be compared to the corresponding case in which the total opinion distribution is randomly given to each community. Thus, eg. if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion 0 communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$. This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\ 

@@

The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced initial opinions, if the opinions are randomly distributed. This is a result that was already found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg. Figure \ref{ev_op_20_80}. If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion 0, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case. This makes sense, since the single opinion communities are more or less isolated from the other communities. This has as consequence that the other (mixed opinion) communities live under the assumption that the opinion distribution is $50/50$. They don't really feel that opinion 0 is a majority opinion, since they are isolated from these single opinion 0 communities. In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer stand. This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present. The community distributed case behaves equivalent to the random distributed case. In the low modularity network, the community structure is not clearly present, but becomes blended. The different communities talk a lot to each other and there is a lot of communication between them. As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion 0 in the same way as if the opinions would be randomly distributed over all the communities.

@@

Lets first discuss the results for the REC method. The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0, with consequently an increase in more mixed neighborhoods. The number of neighborhoods where all the nodes carry opinion 0 thus decreases in favor of neighborhoods where we find both nodes with opinion 0 and nodes with opinion 1. Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion 0 communities, reduces and that the opinions become more mixed. The fraction of opinion 0 remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion 0 see a decrease in the nodes with opinion 0 and become more mixed. We can also observe a very small increase in the number of echo chambers around the minority opinion 1, which indicates the fact that this minority opinion starts to group more together. The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion 0 and opinion 1. Again, the opinion fractions remain constant over time, but are shifted from a random distribution to a more clustered situation. The increase remains however smaller than two, which indicates that there are less than twice as much echo chambers at $t=500$ than at $t=0$.\\
The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion 0, which could be observed for the community distributed, high modularity case. This is due to the fact that the number of echo chambers around opinion 0 at $t=0$, originating from the fraction of single opinion 0 communities, is a lot smaller for the low modularity network than for the high modularity network. The reason for this is the much more blended community structure for the low modularity network. We can still observe a very small increase in the number of echo chambers around opinion 1. This increase in the formation of echo chambers around the minority opinion 1 seems however hampered compared to the randomly distributed case. Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80} (see Figure \ref{sbm}). \\
\newline    % make link to opinion fraction 

@@

Lets first focus on the high modularity network. The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (note that these results were obtained for an initial $20/80$ opinion distribution). We can observe a strong increase in the formation of echo chambers around the majority opinion 0, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion 1. Remember from Figure \ref{high_mod_pr} that the PR method leads to an increase in the majority opinion 0. If we, however, make a fraction equal to $0.1$ of the communities single opinion 0 communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion 0 is strongly hampered. This hampering is partly an artifact of the way the opinions are distributed. If we give entire communities only opinion 0 while others get a $50/50$ distribution, we have a large number of echo chambers around opinion 0 at $t=0$ due to these single opinion communities. Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case. If we, however, would look at the absolute number of echo chambers around opinion 0 at $t=500$, we would see that this number is lower for the community distributed case than for the randomly distributed case. This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0. % share data??

@@

>>>>

As said before, this thesis will try to investigate the interplay between the social network structure and the individual resistance to change in the evolution of two competing opinions. On one hand, we want to determine the influence of different network structures on the formation and evolution of opinions, whereas, on the other hand, we would like to investigate the effect of stubborn agents in the network. Since this thesis deals with opinion dynamics on social media, the effect of filtering algorithms must also be included.\\ 
\newline
In one of their previous papers (\textit{"Modelling opinion dynamics in the age of algorithmic personalization"}, see reference \cite{Perra2019}) prof. Nicola Perra and prof. Luis E.C. Rocha investigated both the effect of different filtering algorithms and different network topologies on the formation and evolution of two competing opinions. They also considered the effect of nudging (meaning that one opinion is pushed to all agents in the network).\\ Their main findings were that the algorithmic filtering could not break the status quo when the prevalence of opinions was equally distributed in the population and that topological correlations (such as high clustering coefficient) could result in the formation of echo chambers \cite{Perra2019}. On the other hand, a more heterogeneous contact pattern could hamper this formation of echo chambers \cite{Perra2019}. If the two opinions were, initially, not equally distributed, one of the filtering algorithms (namely semantic filtering) caused a further increase in the predominant opinion \cite{Perra2019}. In the case of nudging, they found that the opinions in the population moved to the nudged opinion relatively fast, even in the case of small nudging \cite{Perra2019}.\\
They, however, did not investigate the effect of stubborn agents in the network, nor did they investigate networks with a community structure. This thesis will build on their model and incorporate these new features. We will compare networks that have a community structure and relatively low clustering to networks without community structure, but with a high clustering coefficient (a detailed description of concepts like clustering coefficient and community structure will be given in Chapter \ref{chap2}, Section \ref{netDef}). We want to investigate whether networks with community structure are able to form echo chambers, such as one can observe in networks with a high clustering coefficient and, if so, if one can, for example, observe convergence to one opinion inside the communities. We will also pay some attention to networks that have both a high clustering coefficient and a community structure. Can we observe echo chambers in those kinds of networks? If yes, can we determine which property might be the main driver for the formation of echo chambers: high clustering or community structure? \\
% ALSO mention real-world network??
\newline
On the other hand, we want to determine the impact of stubborn agents on the formation and evolution of opinions in the network. Some interesting questions that may come to mind are: Is there an interplay between network structure and stubbornness? What about the possible interplay between the filtering algorithms and the introduction of stubborn people? Does the distribution of stubborn people play a role (eg. randomly distributed versus clustered together inside communities)? \\
% partly rewrite this according to results? Convergence to one opinion in communities --> discussed? Clustered stubborn people --> discussed? Include clustered opinions? Include real-world networks?

@@

Lets look a bit closer to the relation between the formation of echo chambers and the community structure, which is quantified by the modularity $Q$. In Figure \ref{echo_vs_mod}, the size of the echo chambers versus the modularity is depicted. The $y$-axis represents the average of the normalized fraction of nodes with all neighbors carrying opinion 0 and the normalized fraction of nodes with all neighbors carrying opinion 1 (with normalized we mean that the computed value for the fraction of nodes with all neighbors having opinion 0 or all neighbors having opinion 1 at $t=500$ is divided by the corresponding value at $t=0$). This is a measure for the formation of echo chambers (opinion clusters) in the system. The data points represent the value of the echo chamber size in networks with a different modularity. The initial opinion distribution was always $50/50$ an was, initially, randomly distributed over the network. \\

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_PR_20-80_8x7.png}
    \caption{PR}
    \label{PR_all_frac_stubb_20-80}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chamber_vs_all_frac_stubb_REC_20-80_8x7.png}
    \caption{REC}
    \label{REC_all_frac_stubb_20-80}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $20/80$ opinion distribution. \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion 0 and the number of echo chambers around opinion 1) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_all_frac_stubb_PR-REC_20-80}
\end{figure}
\noindent
We will now discuss the results for the evolution of the opinions in the system in case of an unbalanced initial opinion distribution and making all the nodes partly stubborn. Like before, we look at four different network models and for each model we compare the results for three different values of the stubbornness parameter $r$ ($r=0$, $r=0.4$ and $r=0.8$). Two filtering algorithms are investigated, the REC filtering algorithm and the PR filtering algorithm.\\
\newline
Figure \ref{ev_op_20_80_all_frac_stubb_REC} gives the results for the fraction of nodes with opinion 0 versus time in case of the REC filtering algorithm. For none of the network models, nor the values of the stubbornness parameter $r$, the initial prevalence of the opinions can be broken. This result is analogue to the result for an initial $20/80$ opinion distribution, combined with the REC filtering algorithm, in the case of introducing a fraction of completely stubborn people (see Figure \ref{ev_op_20_80_stubb_REC}).

@@

In this subsection, we will look at the stochastic block model (SBM) and we will give a fraction of the communities a single opinion, namely opinion 0. The remaining communities are given a $50/50$ opinion distribution. This situation may, in an idealized way, represent the situation in which you have groups of people following the same idea together with groups where the two ideas coexist. The total opinion distribution over the entire network is no longer $50/50$, but instead opinion 0 becomes the majority opinion (hence, opinion 1 becomes the minority opinion). The higher the fraction of single opinion 0 communities, the bigger the difference between this majority and minority opinion becomes. The results will always be compared to the corresponding case in which the total opinion distribution is randomly distributed over each community. Thus, eg. if we have 10 communities of size 100 and we make a fraction equal to $0.1$ of those communities single opinion 0 communities while the others get a $50/50$ distribution, the total initial opinion distribution would be $55/45$. This situation would then be compared to the case in which this $55/45$ distribution is randomly distributed over each community.\\ 

@@

The PR filtering algorithm leads to an increase of the majority opinion (and thus a decrease of the minority opinion) in case of unbalanced initial opinions, if the opinions are randomly distributed. This is a result that was already found in Section \ref{netw_struc}, Subsection \ref{20-80}, see eg. Figure \ref{ev_op_20_80}. If a fraction equal to $0.1$ of the communities is only given one opinion, namely opinion 0, while the remaining communities get a $50/50$ opinion distribution, we can see from Figure \ref{high_mod_pr} that, in case of the high modularity network, the increase in the majority opinion is hampered compared to the randomly distributed case. This makes sense, since the single opinion communities are more or less isolated from the other communities. This has as consequence that the other (mixed opinion) communities live under the assumption that the opinion distribution is $50/50$. They don't really feel that opinion 0 is a majority opinion, since they are isolated from these single opinion 0 communities. In case of the low modularity network, depicted in Figure \ref{low_mod_pr}, this isolation argument of the different communities no longer holds. This is also clearly visible in the results since the hampering in the increase of the majority opinion is no longer present. The community distributed case behaves equivalent to the random distributed case. In the low modularity network, the community structure is not clearly present, but becomes blended. The different communities talk a lot to each other and there is a lot of communication between them. As a consequence, the mixed opinion communities clearly feel the presence of the single opinion communities and hence feel the majority of opinion 0 in the same way as if the opinions would be randomly distributed over all the communities.

@@

Lets first discuss the results for the REC method. The community distributed case for the high modularity network shows a decrease in the formation of echo chambers around the majority opinion 0 (see the leftmost region on the $x$-axis, where the blue line drops under $1$), with as result an increase in more mixed neighborhoods (see the bump in the blue line). The number of neighborhoods where all the nodes carry opinion 0 thus decreases in favor of neighborhoods where we find both nodes with opinion 0 and nodes with opinion 1. Even though there is only a minor communication between the communities, this limited communication still has the effect that the high number of echo chambers at $t=0$, due to the fraction of single opinion 0 communities, reduces and that the opinions become more mixed. The fraction of opinion 0 remains constant over time when using the REC filtering algorithm, see Figure \ref{high_mod_rec}, however, the communities with the single opinion 0 see a decrease in the nodes with opinion 0 and become more mixed. We can also observe a very small increase in the number of echo chambers around the minority opinion 1, which indicates the fact that this minority opinion starts to group more together. The randomly distributed case for the high modularity network shows an increase in both the echo chambers around opinion 0 and opinion 1. Again, the opinion fractions remain constant over time, but are shifted from a random distribution to a more clustered situation. The increase remains however smaller than two, which indicates that there are less than twice as much echo chambers at $t=500$ than at $t=0$.\\
The community distributed case for the low modularity network does not really show a decrease in the formation of echo chambers around opinion 0, which could be observed for the community distributed, high modularity case. This is due to the fact that the number of echo chambers around opinion 0 at $t=0$, originating from the fraction of single opinion 0 communities, is a lot smaller for the low modularity network than for the high modularity network. The reason for this is the much more blended community structure for the low modularity network. We can still observe a very small increase in the number of echo chambers around opinion 1. This increase in the formation of echo chambers around the minority opinion 1 seems however hampered compared to the randomly distributed case. Remember that the increase in the number of echo chambers around the minority opinion for the stochastic block model in combination with the REC filtering algorithm for a randomly distributed, unbalanced initial opinion distribution was already observed in Section \ref{netw_struc}, Subsection \ref{20-80}, see Figure \ref{sbm}. \\
\noindent    % make link to opinion fraction 

@@

Lets first focus on the high modularity network. The randomly $55/45$ distributed case strongly resembles the results for the stochastic block model in combination with the PR filtering algorithm obtained in Section \ref{netw_struc}, Subsection \ref{20-80}, Figure \ref{sbm} (note that these results were obtained for an initial $20/80$ opinion distribution). We can observe a strong increase in the formation of echo chambers around the majority opinion 0, with a corresponding decrease in all the neighborhoods that have a fraction of nodes with the minority opinion 1. Remember from Figure \ref{high_mod_pr} that the PR method leads to an increase in the majority opinion 0. If we, however, make a fraction equal to $0.1$ of the communities single opinion 0 communities while giving the remaining ones a $50/50$ opinion distribution, we can see that this increase in the formation of echo chambers around the majority opinion 0 is strongly hampered. This hampering is partly an artifact of the way the opinions are distributed. If we give entire communities only opinion 0 while others get a $50/50$ distribution, we have a large number of echo chambers around opinion 0 at $t=0$ due to these single opinion communities. Moreover, the number of opinion 0 echo chambers at $t=0$ is higher than in the corresponding randomly distributed case. If we, however, look at the absolute number of echo chambers around opinion 0 at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case. This shows that the observed hampering of the formation of echo chambers around the majority opinion 0 in the community distributed case compared to the randomly distributed case is a result of more than just an increase in the initial number of echo chambers around this majority opinion 0. Beside this hampering of echo chambers around the majority opinion 0, the community distributed case also shows an increase of echo chambers around the minority opinion 1. Remember from Figure \ref{high_mod_pr} that the high modularity, community distributed case hampers the increase of the majority opinion 0 (and thus the decrease of the minority opinion 1). Hence, a significant fraction of minority opinion nodes (around $30 \%$) is able to survive and they group together in opinion clusters.\\  % share data??
The low modularity network does not show such a discrepancy between the formation of echo chambers in the community distributed and the random distributed cases as could be observed for the high modularity network. Moreover the two cases show qualitatively the same behavior in the formation of echo chambers, with an increase in the number of echo chambers around the majority opinion 0 and a corresponding decrease in all the neighborhoods where a fraction of the nodes carry the minority opinion 1. From Figure \ref{low_mod_pr}, it was already made clear that, in case of the low modularity network, both the random distributed and the community distributed case lead to a significant decrease of this minority opinion (when using the PR filtering algorithm). This decrease of the minority opinion is now also visible in the decrease of neighborhoods where a fraction of the nodes carry this minority opinion. The quantitative difference between the community distributed case and the randomly distributed case (smaller formation of echo chambers around the majority opinion for the former compared to those for the latter) lies partly in the slightly higher number of initial echo chambers around the majority opinion for the community distributed case. However, as for the high modularity case, if we look at the absolute number of echo chambers around the majority opinion at $t=500$, we see that this number is lower for the community distributed case than for the randomly distributed case. % make conclusion about this ? + make table with data??

@@

Figure \ref{echo_05_commOp0_other_50-50} shows a similar plot as Figure \ref{echo_01_commOp0_other_50-50}, but now a fraction equal to $0.5$ of the communities is only given opinion 0, while the others get a $50/50$ distribution. The results look more or less the same as for the case where a fraction equal to $0.1$ of the communities is only given opinion 0 (Figure \ref{echo_01_commOp0_other_50-50}).\\
The REC filtering algorithm, depicted in Figure \ref{rec_comm05_echo}, again shows a decrease in the number of echo chambers around the majority opinion 0 with a corresponding increase in more mixed opinion neighborhoods for the community distributed, high modularity case. A difference with the case where a fraction equal to $0.1$ of the communities are single opinion 0 communities, is the fact that we no longer see an increase in the formation of echo chambers around the minority opinion 1, but instead we observe a decrease in the number of echo chambers around the minority opinion. For the REC filtering algorithm, the initial opinion distribution remains unchanged, but we see that the opinions reorganize in such a way that the final situation mostly has mixed opinion neighborhoods with the majority of the nodes carrying the majority opinion 0 and a minority of the nodes carrying the minority opinion 1. \\
The low modularity, community distributed case seems to wiggle around $1$. If we, however, look closely, we can see that the results qualitatively resemble those for the high modularity, community distributed case (albeit with quantitative values that are a lot smaller). We can observe a small decrease in the formation of echo chambers around the majority opinion 0, with a small increase in more mixed opinion neighborhoods as result. We can also observe a small decrease in the number of echo chambers around the minority opinion 0. The REC filtering algorithm seems to force the system towards a more uniformly mixed opinion distribution in the case of a fraction of $0.5$ of the communities being single opinion 0 communities, while the others get a $50/50$ opinion distribution.\\
The randomly distributed cases for both the high and low modularity strongly resemble the results for an initial, randomly distributed $20/80$ opinion distribution depicted in Figure \ref{sbm}, see Section \ref{netw_struc}, Subsection \ref{20-80}. They show an increase in the formation of echo chambers around the minority opinion, indicating the clustering of the nodes that carry this opinion.\\
Results for the PR filtering algorithm also strongly resemble those depicted in Figure \ref{pr_comm_echo}, which were obtained for a fraction equal to $0.1$ of single opinion 0 communities. The randomly distributed cases for both the high and low modularity are quite similar to those depicted in Figure \ref{sbm} for an initial, random $20/80$ opinion distribution (see Section \ref{netw_struc}, Subsection \ref{20-80}). They show a strong increase in the formation of echo chambers around the majority opinion 0.\\
The low modularity, community distributed case shows a small hampering in the increase of echo chambers around the majority opinion 0, which was also observed in Figure \ref{pr_comm_echo}.\\
The high modularity, community distributed case shows an even stronger hampering in the increase of echo chambers around the majority opinion 0. This was also observed in Figure \ref{pr_comm_echo} and is partly due to the fact that, for a high modularity network, giving a fraction of the communities a single opinion, namely opinion 0, hampers the increase in this majority opinion compared to the randomly distributed case (see Figure \ref{high_mod_pr} and Figure \ref{pr_comm}). The increase in the number of echo chambers around the minority opinion 0 that was visible in Figure \ref{pr_comm_echo} is no longer present.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_REC_8x7.png}
    \caption{REC}
    \label{rec_comm05_echo}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{echo_chambers_SBM_comm05_distr_PR_8x7.png}
    \caption{PR}
    \label{pr_comm05_echo}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution. The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity and a low modularity network. Results for the REC and PR filtering algorithm.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. A fraction equal to $0.5$ of the communities is given a single opinion, namely opinion 0, while the remaining communities get an initial $50/50$ opinion distribution. The community distributed case is compared to the corresponding randomly distributed case ($75/25$) for both a high modularity ($p_{\text{cl}} = 0.1$ and $p_{\text{add}} = 0.001$) and a low modularity ($p_{\text{cl}} = 0.03$ and $p_{\text{add}} = 0.008$) network, both with $N=10^3$ and $\left<k\right> \sim 10$. \textbf{(a)} Results for the REC filtering algorithms; \textbf{(b)} Results for the PR filtering algorithm. Each plot is the result of $10^2$ independent simulations.}
\label{echo_05_commOp0_other_50-50}
\end{figure}
